\batcave: Calling somatic mutations with a tumor- and site-specific prior

### 2019-12-16 ResponseToReviews.tex

\noindent Sincerely,\\
\includegraphics{signature}\\
Ryan Gutenkunst, PhD\\
Associate Professor and Associate Department Head\\
Department of Molecular and Cellular Biology\\
University of Arizona

\section*{Associate Editor}

Stressing on some of the points raised by the reviewers, I would suggest you include more details about the assumption and limitations of BATCAVE. In particular, the possibility of changes of mutational signatures over time can be quite striking in some cases, especially if the original tumour is driven by very specific mutagens like UV light or tobacco smoke. Ideally, the software should try to correct for these cases or at least warn the user about a possible shift in signatures/contexts.
\begin{response}
\RNG{Correcting for context shifts beyond the scope of the current manuscript. But warning about it is a very good idea. Simple check, do a goodness-of-fit test on called the low-confidence mutations. What is the p-value for their profile versus the high-confidence? (A $\chi^2$ test may not be appropriate here, since many entries may have small numbers. This preprint suggests potential improvements: https://arxiv.org/abs/1711.05524.)}
\BKM{This looks quite easy to implement. Testing is another question. One thing we don't have right now is simulations where the signature changes appreciably, but we do have the infrastructure to do it.}
\end{response}

Related to this, there is probably a minimum number of high-confidence mutations that are required by BATCAVE to produce reliable results. As with the previous point, ideally the batcaver should address or warn the user if the assumptions are violated.
\begin{response}
Our simulation studies suggest that a few hundred high-confidence mutations are enough to reliably infer even a sharply peaked mutation profile (Fig. 3).
We have modified \texttt{batcaver} so that it now emits a warning if there are fewer than 500 high-confidence mutations in a sample.
\end{response}

### 2019-12-16 NAR_DRAFT.tex
Response to prior convergence question.

To improve variant identification, the context-dependent prior probability of mutation must converge to an accurate representation of the data generating distribution within the set of high-confidence mutations.
When applied to simulated data, the prior converged within a few hundred mutations (Fig.~\ref{NAR-kl_fig}), \bkmcomment{and the \texttt{batcaver} package emits a warning if there are fewer than 500 high-confidence mutations.}.
For comparison, in our simulated data sets the number of high-confidence mutations ranged between 1,500 and 5,000, and in the real AML we test on it is over 17,000~\cite{Griffith2015}.

>>>>

To improve variant identification, the context-dependent prior probability of mutation must converge to an accurate representation of the data generating distribution within the set of high-confidence mutations.
When applied to simulated data, the prior converged within a few hundred mutations (Fig.~\ref{NAR-kl_fig}), and the \texttt{batcaver} package emits a warning if there are fewer than 500 high-confidence mutations.
For comparison, in our simulated data sets the number of high-confidence mutations ranged between 1,500 and 5,000, and in the real AML we test on it is over 17,000~\cite{Griffith2015}.
